### 2022-07-23 paper/localpsf.tex

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\spatialmean(x')$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

@@

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla\cdot\bm{\sigma}_{\bm{u}}
	&=\rho\bm{g}\,\,\,\text{ in }\Omega, \label{Stokeseqn:1}\\
	\nabla\cdot\bm{u}
	&=0\,\,\,\,\,\,\text{ in }\Omega, \label{Stokeseqn:2}\\
	\bm{\sigma}_{\bm{u}}\bm{n}
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(m\right)\bm{u}\right)
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $m$ is the logarithmic
basal sliding parameter field, $\rho\bm{g}$ the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is a tangential projection operator. 
We employ Glen's flow law
(CITE GLEN) %\cite{glen1955},
which is a constitutive law for ice that relates the stress
tensor $\bm{\sigma}_{\bm{u}}$ and the strain rate tensor
$\bm{\dot{\varepsilon}}_{\bm{u}}= \frac{1}{2}\left(
\bm{\nabla}\bm{u}+\bm{\nabla}\bm{u}^{\top} \right)$, $\bm{\sigma}_{\bm{u}}= 2\eta
\bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.

\begin{figure}[H]
	\begin{subfigure}{0.32\textwidth}
    \includegraphics[scale=0.12]{meshHeight_view0_edges_withColorBar.png}	
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
    \includegraphics[scale=0.12]{trueVelocity_glyphs_withColorBar.png}
    \end{subfigure}
    \begin{subfigure}{0.32\textwidth}
    \includegraphics[scale=0.162]{mtrue_withColorBar.png}
    \end{subfigure}
\end{figure}

\begin{table}[H]
	\begin{center}
		\begingroup
		\setlength{\tabcolsep}{4pt}
		\renewcommand{\arraystretch}{1.25}
		\begin{tabular}{c| c c c | c c c | c c c}
			&  \multicolumn{3}{|c|}{PSF} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
			\hline
			Iter & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ \\
			\verb~0~ &
			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
			\verb~1~ & \verb~4~ & \verb~1.95e+7~ \\
			\verb~1~ &
			\verb~2~ & \verb~6~ & \verb~5.76e+6~ &
			\verb~2~ & \verb~6~ & \verb~7.90e+6~ &
			\verb~2~ & \verb~6~ & \verb~5.76e+6~ \\
			\verb~2~ &
			\verb~4~ & \verb~10~ & \verb~2.36e+6~ &
			\verb~5~ & \verb~12~ & \verb~3.87e+6~ &
			\verb~4~ & \verb~10~ & \verb~2.36e+6~ \\
			\verb~3~ &
			\verb~2~ & \verb~6+22~ & \verb~5.79e+5~ &
			\verb~13~ & \verb~28~ & \verb~1.62e+6~ &
			\verb~13~ & \verb~28~ & \verb~5.79e+5~ \\
			\verb~4~ &
			\verb~5~ & \verb~12~ & \verb~5.61e+4~ &
			\verb~42~ & \verb~86~ & \verb~4.78e+5~ &
			\verb~33~ & \verb~68~ & \verb~9.05e+4~ \\
			\verb~5~ &
			\verb~10~ & \verb~22~ & \verb~3.45e+3~ &
			\verb~84~ & \verb~170~ & \verb~7.72e+4~ &
			\verb~57~ & \verb~116~ & \verb~6.21e+3~ \\
			\verb~6~ &
			\verb~14~ & \verb~30~ & \verb~2.66e+1~ &
			\verb~125~ & \verb~252~ & \verb~5.04e+3~ &
			\verb~76~ & \verb~154~ & \verb~1.05e+2~ \\
			\verb~7~ &
			\verb~0~ & \verb~2~ & \verb~4.09e-2~ &
			\verb~194~ & \verb~390~ & \verb~7.91e+1~ &
			\verb~117~ & \verb~236~ & \verb~2.29e-1~ \\
			\verb~8~ & 
			\verb~---~ & \verb~---~ & \verb~---~ &
			\verb~0~ & \verb~2~ & \verb~1.60e-1~ &
			\verb~0~ & \verb~2~ & \verb~4.08e-2~ \\
			\hline
			Total & 
			\verb~38~ & \verb~114~ & \verb~---~ &
			\verb~466~ & \verb~950~ & \verb~---~ &
			\verb~303~ & \verb~624~ & \verb~---~ \\
		\end{tabular}
		\endgroup
	\end{center}
\end{table}

\begin{figure}[H]
	\begin{subfigure}{0.225\textwidth}
		\includegraphics[scale=0.08]{m1_psf.png}	
	\end{subfigure}
	\begin{subfigure}{0.225\textwidth}
		\includegraphics[scale=0.08]{m3_psf.png}
	\end{subfigure}
	\begin{subfigure}{0.225\textwidth}
		\includegraphics[scale=0.08]{m4_psf.png}
	\end{subfigure}
	\begin{subfigure}{0.225\textwidth}
	\includegraphics[scale=0.08]{m6_psf.png}
\end{subfigure}
\begin{subfigure}{0.05\textwidth}
	\includegraphics[scale=0.1]{mColorBar.png}
\end{subfigure}
\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).} 
\end{figure}

>>>>

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\spatialmean(x')$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

@@

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla\cdot\bm{\sigma}_{\bm{u}}
	&=\rho\bm{g}\,\,\,\text{ in }\Omega, \label{Stokeseqn:1}\\
	\nabla\cdot\bm{u}
	&=0\,\,\,\,\,\,\text{ in }\Omega, \label{Stokeseqn:2}\\
	\bm{\sigma}_{\bm{u}}\bm{n}
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(\beta\right)\bm{u}\right)
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\omega\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $\beta$ is the logarithmic
basal sliding parameter field, $\rho\bm{g}$ the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is a tangential projection operator. 
We employ Glen's flow law
(CITE GLEN) %\cite{glen1955},
which is a constitutive law for ice that relates the stress
tensor $\bm{\sigma}_{\bm{u}}$ and the strain rate tensor
$\bm{\dot{\varepsilon}}_{\bm{u}}= \frac{1}{2}\left(
\bm{\nabla}\bm{u}+\bm{\nabla}\bm{u}^{\top} \right)$, $\bm{\sigma}_{\bm{u}}= 2\eta
\bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model. The domain geometry $\Omega$ is shown in Figure \ref{fig:iceMountainMesh}.

The inverse problem that we seek to solve is

\begin{align*}
\min_{\beta}J(\beta)&:=\frac{1}{2}\int_{\Gamma_{t}}||\bm{T}(\bm{u}-\bm{u}_{d})||_{2}^{2}\,dS+\frac{1}{2}\int_{\Omega}\beta\mathcal{L}^{2}\beta\,dx,\\
\text{where }\bm{u}& \text{ solves \ref{Stokeseqn}}.
\end{align*}
Here $\mathcal{L}$ is a regularizing invertible elliptic differential operator, where $s=\mathcal{L}^{-1}u$ if and only if

\begin{align*}
-\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Gamma_{b},\\
\gamma \bm{\nabla}u\cdot\bm{n}+\mu\,u&=0\,\,\,\text{ on }\partial\Gamma_{b}.
\end{align*}
The observation data $\bm{u}_{d}(\bm{x})$ is generated synthetically, as follows, we propose a true log basal friction field $\beta_{\text{true}}(\bm{x})$, solve equation \ref{Stokeseqn} to obtain $\bm{u}_{\text{true}}(\bm{x})$, we then add random relative noise to obtain the noisy data $\bm{u}_{d}$

\begin{align*}
\left(\bm{u}_{d}(\bm{x})\right)_{i}&=\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}+\varepsilon\, \xi_{i}(\bm{x})|\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}|\,\,\,i=1,\dots,d,
\end{align*}
where $\xi_{i}(\bm{x})$ are standard normal random numbers and $\varepsilon>0$ sets the noise level. The inverse problem is solved with an inexact-Newton CG scheme, wherein each Hessian $\mathcal{A}$ action, requires the linearized PDE solutions, and so we apply the psf framework to generate Hessian preconditioners and compare the performance with other techniques, such as preconditioning by the regularization operator.

\begin{figure}[H]
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
	\end{subfigure}
	\begin{subfigure}{0.38\textwidth}
		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
	\end{subfigure}
\caption{TO DO: description of ice mountain}
\label{fig:iceMountainMesh}
\end{figure}

\begin{figure}[H]
    \begin{subfigure}{0.49\textwidth}
		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
	\end{subfigure}
    \begin{subfigure}{0.49\textwidth}
    	\includegraphics[scale=0.18]{trueVelocity_glyphs_withColorBar.png}
    \end{subfigure}
\end{figure}

\begin{table}[H]
	\begin{center}
		\begingroup
		\setlength{\tabcolsep}{3pt}
		\renewcommand{\arraystretch}{1.25}
		\begin{tabular}{c| c c c | c c c | c c c}
			&  \multicolumn{3}{|c|}{PSF} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
			\hline
			Iter & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ \\
			\verb~0~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
			\verb~1~ &
			\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
			\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
			\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
			\verb~2~ &
			\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
			\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
			\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
			\verb~3~ &
			\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
			\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
			\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
			\verb~4~ &
			\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
			\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
			\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
			\verb~5~ &
			\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
			\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
			\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
			\verb~6~ &
			\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
			\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
			\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
			\verb~7~ &
			\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
			\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
			\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
			\verb~8~ & 
			\verb~---~ & \verb~---~ & \verb~---~ &
			\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
			\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
			\hline
			Total & 
			\verb~38~ & \verb~114~ & \verb~---~ &
			\verb~466~ & \verb~950~ & \verb~---~ &
			\verb~303~ & \verb~624~ & \verb~---~ \\
		\end{tabular}
		\endgroup
	\end{center}
\end{table}

\begin{figure}[H]
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m1_psf.png}	
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m3_psf.png}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m4_psf.png}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
	\includegraphics[scale=0.1]{m6_psf.png}
\end{subfigure}
%\begin{subfigure}{0.05\textwidth}
%	\includegraphics[scale=0.1]{mColorBar.png}
%\end{subfigure}
\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).} 
\end{figure}
